---
title: "RDS-assignment"
author: "B224911"
date: "2022-11-11"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, results="hide")
library(pheatmap)
library(dplyr)
library(lintr)
lint("./RDS_script.R")
```
## R Markdown for the assignment of RDS
##Processing the data and make the heatmap
#because view can not be used in knit, so it not show in report

step1.
#Load the data into the environment 
```{r OrignalData}
gene_annot <- read.csv("gene_annotation.csv")  #  read gene annotation
sample_annot <- read.csv("sample_annotation.csv")  #  read sample annotation
ori_data <- read.csv("data_all.csv")  # read numerical unprocessed data
# read the list of genes to use for plotting
genelist <- read.table("genelist_15.txt", header = TRUE, sep = "\t")
# Check the data
```

step2.
#convert the data_all to log scaled data.
#Select the required data for the plot.
#Each gene should only appear once on the final plot.

```{r LogAndDup}
log_data <- log(ori_data[, -1])
# Select the required data for the plot
# Each gene should only appear once on the final plot
dup_list <- duplicated(genelist$x)
nodup_list <- sort(genelist[!dup_list, ])
nodup_list  # check the unduplicated gene list
class(nodup_list)
# Select the unduplicated unprocessed data
select_data <- log_data[nodup_list, ]
```

step3.
#In the plot annotate each gene row with the type of gene it is (XA, XB or XC) 
#and annotate each sample with the treatment group that the sample was derived from.
#Before plotting the samples, rename the gene names with the “LongName” from the gene annotation table.

```{r annotaiton}
g_annot <- gene_annot[nodup_list, ]
identical(rownames(g_annot), rownames(select_data))
# Build the row and column annotation index
# Select one colume which is genetype column of 'g_annot' as the row annotation
# but it is not the dataframe type, so change its tpye
# and rename the rownames as same as the genenames(“LongName”) of select_data
annot_row <- (g_annot[, 3])
class(annot_row)
annot_row <- data.frame(g_annot[, 3])
# Create a nicer name for genetype
annot_row <- rename(annot_row,  "GeneType" = "g_annot...3.")
rownames(annot_row) <- g_annot$LongName

# Select the treatment group column of sample_annot as the column annotation
# and deal with the 'annot_col' in the same way
annot_col <- data.frame(sample_annot[, 3])
# create a nicer name for treatmentgroup
annot_col <- rename(annot_col, "TreatmentGroup" = "sample_annot...3.")
rownames(annot_col) <- sample_annot[, 2]
# Rename the gene names with the “LongName”
rownames(select_data) <- g_annot$LongName

```


step4.
#make the heatmap
```{r heatmap}
pheatmap(
  select_data,
  scale = 'row',
  annotation_col = annot_col,
  annotation_row = annot_row,
  cluster_col = TRUE,
  cluster_row = TRUE,  # Both cluster
  main = 'Heatmap of genes and the samples are clustered',
  legend = TRUE,
  fontsize_row = 6,
  fontsize_col = 12
  )
# Plot2
pheatmap(
  select_data,
  scale = 'row',
  annotation_col = annot_col,
  annotation_row = annot_row,
  cluster_col = FALSE,
  cluster_row = TRUE,  # Only gene cluster
  main = 'Heatmap of only the genes are clustered',
  legend = TRUE,
  fontsize_row = 6,
  fontsize_col = 12
  )
```
